Meta-analysis

### 2020-06-24 04_MetaAnalysis.Rmd
Cleaning, editing, colours, grammar, spelling, headings, etc.

# Meta-analysis
## Grain yield meta-analysis

@@

First, how well the data is compared across the trial years and trials.

```{r table_of_nTreatments}
kable(
   table(slimmer_PM_dat$spray_management, slimmer_PM_dat$year),
   align = rep('c', 8),
   caption = "Which treatments and how many treatments are represented in each year"
) %>%
   kable_styling(
      "striped",
      fixed_thead = TRUE,
      full_width = FALSE,
      position = "center"
   )
```

Before analysis let's have a look at the trimmed down modified data.

Treatments Late_plus and early don't have very good comparison to other treatments.

@@

```{r treatment_means_plot}
# class spray_management as a factor and reorder them for the plot
slimmer_PM_dat$spray_management <-
   factor(slimmer_PM_dat$spray_management,
          levels(factor(slimmer_PM_dat$spray_management))[rev(c(1, 2, 5, 6, 3, 4))])

slimmer_PM_dat %>%
   ggplot(aes(y = grain_yield.t.ha, x = spray_management)) +
   geom_boxplot() +
   #geom_point(position = "jitter", alpha = 1/5)+
   geom_jitter(width = 0.1, alpha = 1 / 5) +
   labs(x = "Spray management variable",
        y = "Grain yield (t/Ha)",
        title = "Mean grain yield from each treatment \n categorised by spray management scenario") +
   theme(plot.title = element_text(hjust = 0.5)) +
   geom_hline(yintercept = 0, linetype = 2) +
   coord_flip()
```

There seems like little difference between the treatments, with exception to `Late_plus`.
Let's do this plot again, but let's look at the proportional mean difference between the treatments and the no spray control for each study, this should reduce variation in yield due between trials.

```{r plot_yieldProportion}
slimmer_PM_dat$spray_management <-
   factor(slimmer_PM_dat$spray_management, rev(
      c(
         "control",
         "Early",
         "Recommended",
         "Recommended_plus",
         "Late",
         "Late_plus"
      )
   ))

slimmer_PM_dat %>%
   ggplot(aes(y = prop_yield_gain, x = spray_management)) +
   geom_boxplot() +
   geom_jitter(width = 0.1, alpha = 1 / 5) +
   labs(x = "Spray management variable",
        y = "Grain yield (t/Ha)",
        title = "Mean grain yield difference to the control for each treatment \n categorised by spray management scenario") +
   theme(plot.title = element_text(hjust = 0.5)) +
   geom_hline(yintercept = 0, linetype = 2) +
   coord_flip()
```

@@

First lets remove any entries missing variance.

@@

slimmer_PM_dat <- 
   slimmer_PM_dat[!is.na(slimmer_PM_dat$yield_error),]

@@

```{r MSE_2_SVar}
#calculation when using log of response
# slimmer_PM_dat$vi <-
#    slimmer_PM_dat$Y_Msquare / (slimmer_PM_dat$n * slimmer_PM_dat$grain_yield.t.ha ^
#                                   2)
#
slimmer_PM_dat$vi2 <- 
   slimmer_PM_dat$Y_Msquare / 
   slimmer_PM_dat$n
```

@@

```{r StandardisedMeanDifference}
slimmer_PM_dat$pooledSD <- NA
slimmer_PM_dat$vi_C <- NA
for (T_ref in slimmer_PM_dat$trial_ref) {
   # First we need to Calculate the pooled standard deviation
   slimmer_PM_dat[slimmer_PM_dat$trial_ref == T_ref, "pooledSD"] <-
      sqrt(sum((slimmer_PM_dat[slimmer_PM_dat$trial_ref == T_ref, "n"] - 1) * slimmer_PM_dat[slimmer_PM_dat$trial_ref == T_ref, "yield_error"]) /
              (sum(slimmer_PM_dat[slimmer_PM_dat$trial_ref == T_ref, "n"]) - nrow(slimmer_PM_dat[slimmer_PM_dat$trial_ref == T_ref,])))

   # Then Create a column with the value of the mean grain yield of the no spray control for comparisons to the treatments means
   slimmer_PM_dat[slimmer_PM_dat$trial_ref == T_ref, "yi_C"] <-
      mean(slimmer_PM_dat[slimmer_PM_dat$trial_ref == T_ref &
                             slimmer_PM_dat$fungicide_ai == "control",
                          "grain_yield.t.ha"], na.rm = TRUE)

}
# Calculate standardised mean difference
slimmer_PM_dat$grain_SMD <-
   (slimmer_PM_dat$grain_yield.t.ha - slimmer_PM_dat$yi_C) / slimmer_PM_dat$pooledSD 
```

@@

Let's simplify our data.

@@

## Meta-analysis

### metafor package

@@

Standardised mean differences have a bit of a long tail however this is not bad and transformations with log or square root will not work on negative values.

@@

```{r Metafor-analysis}
PM_mv <- rma.mv(
   yi = grain_SMD,
   vi,
   mods = ~ spray_management - 1,
   method = "ML",
   random = ~ spray_management | trial,
   struct = "UN",
   data = dat1
)
summary(PM_mv)
```
In this result we can see that the `Early` treatment is not significantly different to zero, which in this case is the mean of the no spray control. 
However the other treatments are significantly different to zero.
The $Q_M$ [omnibus test](http://www.metafor-project.org/doku.php/tips:models_with_or_without_intercept?s[]=anova) of moderators, shows the moderators significantly influence the model ($Q_M = 24.4769, df = 5, p = 0.0002$) and we can reject the null hypothesis ($H_0 : \beta_1 = \beta_2 = \beta_3 =\beta_4 = 0$) that there is no difference between the moderators [@Viechtbauer2010].
The analysis shows there is still a significant amount of residual heterogeneity ($Q_E = 47501.76, df = 111, p < 0.0001$) not captured by the spray management moderator indicating other possible moderators which might influence grain yield.

@@

```{r rma_noRandomInteractiveTerm}
PM_mv_St <- rma.mv(
   yi = grain_SMD,
   vi, 
   mods = ~ spray_management,
   method = "ML",
   random = list(~ 1|trial),
   data = dat1
)

PM_mv_St1 <- rma.mv(
   yi = grain_SMD,
   vi, 
   mods = ~ spray_management,
   method = "ML",
   random = list(~ 1|location/year/trial_ref/dose/host_genotype/factor(row_spacing)/spray_management),
   data = dat1
)

AIC(PM_mv) # Printing the AIC of the model helps differentiate the models in the anova output
anova(PM_mv, PM_mv_St1) # PM_mv is the better model

```

The simpler `PM_mv_St` model explains significantly less variation as the more complicated model `PM_mv`.
So we will keep the original model `PM_mv`.

The first table in this output shows the tau^2 (variance) of each random effects and the number of occurrences for each treatment in the analysis.
The second table is in two parts(left and right). 
The left part, rho, is the correlation of variation between the specified treatments. 
All comparisons were acceptable except for a comparison between `Early` and `Late_plus`, `0.000` rho. `Early` and `Late_plus` treatments never occurred within the same trial which is indicated by the right side of the table (hence the warning).

The fixed effects, in the last table, showed that yields in single early spray treatments are not significantly different from zero, or the no spray control.

@@

Results show with the exclusion of the no spray control, `Early` applications are significantly worse than all other treatments leading to lower yield estimates. 
`Late_plus` is estimated as saving the most yield compared and is significantly higher than all other treatments except `recommended_plus`.
Let's view these comparisons in a plot.

@@

efficacy <- tbl_df(results_AI)
efficacy$Treatment <-
   factor(c(
      "Early",
      "Recommended",
      "Recommended_plus",
      "Late",
      "Late_plus"
   ))

@@

```{r metafor_plot}
efficacy %>%
   mutate(Treatment = factor(Treatment, levels = rev(c("Early", "Recommended", "Recommended_plus","Late","Late_plus"))))%>%
   ggplot(aes(Treatment, Mean)) +
   geom_hline(
      yintercept = seq(-0.4, 1.8, by = 0.2),
      color = usq_cols("usq charcoal"),
      linetype = 3
   ) +
   geom_hline(yintercept = 0)+
   geom_point(aes(size = 1 / SE), shape = 15) +
   geom_linerange(aes(ymin = CIs_lower, ymax = CI_upper)) +
   coord_flip()

```

Let's look at how well each of the number treatments compare to each other.

@@

We also want to know if disease pressure exerts an effect on the desicion for spray control.
Above we can see spraying at the first sign of powdery mildew or with in `r min(as.numeric(slimmer_PM_dat[slimmer_PM_dat$spray_management == "Late"| slimmer_PM_dat$spray_management == "Late_plus","fungicide_timing_1"]))` - `r max(as.numeric(slimmer_PM_dat[slimmer_PM_dat$spray_management == "Late"|                   slimmer_PM_dat$spray_management == "Late_plus","fungicide_timing_1"]))` days after first sign is effective at mitigating yield loss.
Does this change between years of low and high disease pressure.

@@

These results show that the spray management variable we tested are not significantly better or worse between the two levels of disease pressure.
It also shows that in trials with high disease pressure all treatments produced lower standardised mean differences. 
Indicating that fungicide treatments were less effective at mitigating yeild loss in seasons which developed heavier infestations.  
Finally lets compare the disease pressure model to the orignal model to see which one fits better.

```{r compareDvNDP}
anova(PM_mv, PM_mv_Dp)
```
No significant difference which informs us to use the origninal model `PM_mv` which is the simplest.

@@

We can use the netmeta package to give a graphical representation of the pairwise comparisons.

### netmeta package

@@

The `netmeta` analysis suggests the spray schedule commencing early are no different to any other treatment including the no spray control.
It estimates the mean is very similar to the recommended treatments.
The recommended plus and late_plus treatments show higher mean estimates, however are not significantly different from the early estimate.

>>>>

# Grain yield meta-analysis

@@

First, how well does the data compare across the trial years and trials?

```{r table_of_nTreatments}
# class spray_management as a factor and reorder them for the plot
slimmer_PM_dat$spray_management <-
   factor(slimmer_PM_dat$spray_management, rev(
      c(
         "control",
         "Early",
         "Recommended",
         "Recommended_plus",
         "Late",
         "Late_plus"
      )
   ))

ggplot(slimmer_PM_dat, aes(x = as.factor(year), fill = spray_management)) +
   geom_bar(position = "dodge2") +
   scale_fill_usq(name = "Spray Management") +
   xlab("Year")

kable(
   table(slimmer_PM_dat$spray_management, slimmer_PM_dat$year),
   align = rep('c', 8),
   caption = "Which treatments and how many treatments are represented in each year"
) %>%
   kable_styling(
      "striped",
      fixed_thead = TRUE,
      full_width = FALSE,
      position = "center"
   )
```

Treatments `Late_plus` and `Early` don't have very good comparison to other treatments.

@@

```{r treatment_means_plot}
slimmer_PM_dat %>%
   ggplot(
      aes(
         y = grain_yield.t.ha,
         x = spray_management,
         fill = spray_management,
         colour = spray_management
      )
   ) +
   geom_boxplot(alpha = 0.25) +
   geom_jitter(width = 0.1, alpha = 0.5) +
   labs(x = "Spray management variable",
        y = "Grain yield (t/Ha)",
        title = "Mean grain yield from each treatment \n categorised by spray management scenario") +
   theme(plot.title = element_text(hjust = 0.5)) +
   geom_hline(yintercept = 0, linetype = 2) +
   scale_fill_usq() +
   scale_colour_usq() +
   coord_flip() +
   guides(fill = FALSE, color = FALSE)
```

There seems like little difference between the treatments, with exception of `Late_plus`.
Let's do this plot again, but let's look at the proportional mean difference between the treatments and the no spray control for each study, this should reduce variation in yield due between trials.

```{r plot_yieldProportion}
slimmer_PM_dat %>%
   filter(spray_management != "control") %>% 
   ggplot(
      aes(
         y = prop_yield_gain,
         x = spray_management,
         fill = spray_management,
         colour = spray_management
      )
   ) +
   geom_boxplot(alpha = 0.25) +
   geom_jitter(width = 0.1, alpha = 0.5) +
   labs(x = "Spray management variable",
        y = "Grain yield (t/Ha)",
        title = "Mean grain yield from each treatment \n categorised by spray management scenario") +
   theme(plot.title = element_text(hjust = 0.5)) +
   geom_hline(yintercept = 0, linetype = 2) +
   scale_fill_usq() +
   scale_colour_usq() +
   coord_flip() +
   guides(fill = FALSE, color = FALSE)
```

@@

First let's remove any entries missing variance.

@@

slimmer_PM_dat <-
   slimmer_PM_dat[!is.na(slimmer_PM_dat$yield_error), ]

@@

```{r MSE_2_SVar}
#calculation when using log of response
# slimmer_PM_dat$vi <-
#    slimmer_PM_dat$Y_Msquare / (slimmer_PM_dat$n * slimmer_PM_dat$grain_yield.t.ha ^
#                                   2)

slimmer_PM_dat$vi2 <- 
   slimmer_PM_dat$Y_Msquare / 
   slimmer_PM_dat$n
```

@@

```{r StandardisedMeanDifference}
slimmer_PM_dat$pooledSD <- NA
slimmer_PM_dat$vi_C <- NA
for (i in slimmer_PM_dat$trial_ref) {
   # First we need to Calculate the pooled standard deviation
   slimmer_PM_dat[slimmer_PM_dat$trial_ref == i, "pooledSD"] <-
      sqrt(sum((slimmer_PM_dat[slimmer_PM_dat$trial_ref == i, "n"] - 1) * slimmer_PM_dat[slimmer_PM_dat$trial_ref == i, "yield_error"]) /
              (sum(slimmer_PM_dat[slimmer_PM_dat$trial_ref == i, "n"]) - nrow(slimmer_PM_dat[slimmer_PM_dat$trial_ref == i, ])))

   # Create a column with the value of the mean grain yield of the no spray
   #  control for comparisons to the treatments means
   slimmer_PM_dat[slimmer_PM_dat$trial_ref == i, "yi_C"] <-
      mean(slimmer_PM_dat[slimmer_PM_dat$trial_ref == i &
                             slimmer_PM_dat$fungicide_ai == "control",
                          "grain_yield.t.ha"], na.rm = TRUE)

}

# Calculate standardised mean difference
slimmer_PM_dat$grain_SMD <-
   (slimmer_PM_dat$grain_yield.t.ha - slimmer_PM_dat$yi_C) / slimmer_PM_dat$pooledSD
```

@@

Let's simplify our data by removing the control data, converting disease pressure (`D_pres`) to a factor and selecting only the columns of data necessary for analysis.

@@

## metafor analysis

For the first grain yield meta-analysis, we'll use the `metafor` package [@Viechtbauer2010].

@@

Standardised mean differences have a bit of a long tail however this is not bad and transformations with log or square root will not work on negative values.
So it looks like no transformation is necessary before we start the meta-analysis.

@@

```{r Metafor-analysis}
PM_mv <- rma.mv(
   yi = grain_SMD,
   vi,
   mods = ~ spray_management - 1,
   method = "ML",
   random = ~ spray_management | trial,
   struct = "UN",
   data = dat1
)
summary(PM_mv)
```

In this result we can see that the `Early` treatment is not significantly different from zero, which in this case is the mean of the no spray `control`.
However the other treatments are significantly different from zero.
The $Q_M$ [omnibus test](http://www.metafor-project.org/doku.php/tips:models_with_or_without_intercept?s[]=anova) of moderators, shows the moderators significantly influence the model ($Q_M = 24.4769, df = 5, p = 0.0002$) and we can reject the null hypothesis ($H_0 : \beta_1 = \beta_2 = \beta_3 =\beta_4 = 0$) that there is no difference between the moderators [@Viechtbauer2010].
The analysis shows there is still a significant amount of residual heterogeneity ($Q_E = 47501.76, df = 111, p < 0.0001$) not captured by the spray management moderator indicating other possible moderators which might influence grain yield.

@@

```{r rma_noRandomInteractiveTerm}
PM_mv_St <- rma.mv(
   yi = grain_SMD,
   vi,
   mods = ~ spray_management,
   method = "ML",
   random = list( ~ 1 | trial),
   data = dat1
)

PM_mv_St1 <- rma.mv(
   yi = grain_SMD,
   vi,
   mods = ~ spray_management,
   method = "ML",
   random = list(
      ~ 1 |
         location / year / trial_ref / dose / host_genotype / factor(row_spacing) /
         spray_management
   ),
   data = dat1
)

AIC(PM_mv) # Printing the AIC of the model helps differentiate the models in the anova output
anova(PM_mv, PM_mv_St1) # PM_mv is the better model
```

The simpler `PM_mv_St` model explains significantly less variation than the more complicated model `PM_mv`.
So we will keep the original model `PM_mv`.

The first table in this output shows the tau^2 (variance) of each random effects and the number of occurrences for each treatment in the analysis.
The second table is in two parts (left and right). 
The left part, rho, is the correlation of variation between the specified treatments.
All comparisons were acceptable except for a comparison between `Early` and `Late_plus`, `0.000` rho.
`Early` and `Late_plus` treatments never occurred within the same trial, which is indicated by the right side of the table (hence the warning).

The fixed effects, in the last table, show that yields in single early spray treatments are not significantly different from zero, or the no spray control.

@@

Results show with the exclusion of the no spray `control`, `Early` applications are significantly worse than all other treatments leading to lower yield estimates.
`Late_plus` is estimated to save the most yield compared and is significantly higher than all other treatments except `Recommended_plus`.
Let's look at how well each of the number treatments compare to each other.

@@

efficacy <- as_tibble(results_AI)
efficacy$Treatment <-
   factor(c(
      "Early",
      "Recommended",
      "Recommended_plus",
      "Late",
      "Late_plus"
   ))

@@

Let's view these comparisons in a plot.

```{r metafor_plot}
efficacy %>%
   mutate(Treatment = factor(Treatment, levels = rev(
      c(
         "Early",
         "Recommended",
         "Recommended_plus",
         "Late",
         "Late_plus"
      )
   ))) %>%
   ggplot(aes(Treatment, Mean)) +
   geom_hline(
      yintercept = seq(-0.4, 1.8, by = 0.2),
      color = usq_cols("usq charcoal"),
      linetype = 3
   ) +
   geom_hline(yintercept = 0) +
   geom_point(aes(size = 1 / SE), shape = 15) +
   geom_linerange(aes(ymin = CIs_lower, ymax = CI_upper)) +
   coord_flip()
```

@@

We also want to know if disease pressure exerts an effect on the decision for spray control.
Above we can see spraying at the first sign of powdery mildew or with in `r min(as.numeric(slimmer_PM_dat[slimmer_PM_dat$spray_management == "Late"| slimmer_PM_dat$spray_management == "Late_plus","fungicide_timing_1"]))` - `r max(as.numeric(slimmer_PM_dat[slimmer_PM_dat$spray_management == "Late"|                   slimmer_PM_dat$spray_management == "Late_plus","fungicide_timing_1"]))` days after first sign is effective at mitigating yield loss.
Does this change between years of low and high disease pressure.

@@

These results show that the spray management variable we tested are not significantly better or worse between the two levels of disease pressure.
It also shows that in trials with high disease pressure, all treatments produced lower standardised mean differences. 
Indicating that fungicide treatments were less effective at mitigating yield loss in seasons, which developed more severe infection levels.
Finally let's compare the disease pressure model to the original model to see which one fits better.

```{r compareDvNDP}
anova(PM_mv, PM_mv_Dp)
```

No significant difference which informs us to use the original model `PM_mv`, which is the simplest.

@@

## netmeta analysis

We can use the `netmeta` package to give a graphical representation of the pairwise comparisons.

@@

The `netmeta` analysis suggests the spray schedule commencing early are no different to any other treatment including the no spray `control`.
It estimates the mean is very similar to the recommended treatments.
The `Recommended_plus` and `Late_plus` treatments show higher mean estimates, however are not significantly different from the `Early` estimate.
